---
title: "canid fossil plots"
output: html_document
---

```{r}
library(ggplot2)
canids <- read.csv("canid_range.csv", header = F, as.is = T)

names(canids) <- c("genus", "species", "minage", "maxage")
head(canids)

canid_occ <- ggplot(canids, aes( species, ymin = maxage, ymax=minage, colour = genus))
canid_occ <- canid_occ + geom_linerange()
canid_occ

canid_occ <- canid_occ + theme(legend.position="none")
canid_occ

canid_occ <- canid_occ + coord_flip()
canid_occ

canid_occ <- canid_occ +  theme(axis.text.y = element_text(size=3))
canid_occ

canid_occ <- canid_occ + theme(axis.ticks.y=element_blank())
canid_occ

canid_occ <- canid_occ + scale_y_continuous(limits=c(0, 40), expand = c(0, 0), breaks=c(0, 10, 20, 30, 40))
canid_occ

canid_occ <- canid_occ + labs(title = "Canid Fossil Occurrences", x = "Species", y = "Ma ago") + theme(plot.title = element_text(hjust = 0.5, size=22, face = "bold"), axis.title =element_text(size=20))
canid_occ

ggsave(filename = "canid-occ.pdf", plot = canid_occ)

```

Q: What taxonomic group is displayed?
A: Canids

Q: How many unique fossil species are contained within it?
A: 171
```{r}
length(canids$species)
```

Q: How many unique fossil genera?
A: 58
```{r}
nrow(canids[!duplicated(canids$genus), ])
```
Q: What species has the longest range
A: Cormocyon copei
```{r}
canids[which(
  (canids$maxage - canids$minage) == max(canids$maxage - canids$minage)), 2
  ]
```

## Redo this exercise with another taxon
### Taxon: Equidae

```{r}
Equids <- read.csv("Equid_range.csv", header = F, as.is = T)

names(Equids) <- c("genus", "species", "minage", "maxage")
head(Equids)

Equid_occ <- ggplot(Equids, aes(species, ymin = maxage, 
                                ymax=minage, colour = genus)) + 
  geom_linerange() +
  theme(legend.position="none", axis.text.y = element_text(size = 3),
        axis.ticks.y = element_blank(), plot.title = element_text(hjust = 0.5,
                                                                size=22, 
                                                                face = "bold"),
        axis.title =element_text(size=20)) +
  coord_flip() +
  scale_y_continuous(limits=c(0, 40), 
                     expand = c(0, 0), breaks=c(0, 10, 20, 30, 40)) +
  labs(title = "Equidae Fossil Occurrences", x = "Species", y = "Ma ago")
  
Equid_occ

ggsave(filename = "Equid-occ.pdf", plot = Equid_occ)
```
Q: What taxonomic group is displayed?
A: Equidae

Q: How many unique fossil species are contained within it?
A: 219
```{r}
length(Equids$species)
```

Q: How many unique fossil genera?
A: 43
```{r}
nrow(Equids[!duplicated(Equids$genus), ])
```
Q: What species has the longest range
A: Eurohippus parvulus
```{r}
Equids[which(
  (Equids$maxage - Equids$minage) == max(Equids$maxage - Equids$minage)), 2
  ]
```



# Part 2: Data Frames
## Exploring Data Frames

```{r}
# Challenge 1
df <- data.frame(First_name = c("Clive"),
                 Last_name = c("Lau"),
                 Lucky_number = c(5),
                 stringsAsFactors = FALSE)
df <- rbind(df, list("John", "Snow", 4048))
df <- cbind(df, Coffee_break = c("Y", "Y"))

# Challenge 2
source("swcarpentry-R-challenge2.R")

# Challenge 3
str(gapminder)
# output yields the number of entries and number of variables. It says it is a dataframe with columns "country", "year", "pop", "continent", "lifeExp", and "gdpPercap". 

```

## Subsetting Data

```{r}
# Challenge 1
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

x[c(2, 3, 4)]
x[2:4]
x[-c(1, 5)]

# Challenge 2
x[-which(names(x) == "g")]
names(x) == "g" # returns T/F for each of the named variables in x checking if they are named "g"
which(names(x) == "g") # returns how many of the variables are named "g", in this case, zero.
x[-which(names(x) == "g")] # returns empty vector
# Statments A and C are true

# Challenge 3
x[which(x < 7 & x >4)]

# Challenge 4
m <- matrix(1:18, nrow=3, ncol=6)
print(m)
m[2,c(4,5)]
# answer D will extract values 11 and 14

# Challenge 5
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))
xlist$b[2]

# Challenge 6
mod <- aov(pop ~ lifeExp, data=gapminder)
attributes(mod)
mod$df.residual

# Challenge 7
# Extract observations collected for the year 1957
gapminder[gapminder$year == 1957,]
#Extract all columns except 1 through to 4
gapminder[,-(1:4)]
#Extract the rows where the life expectancy is longer the 80 years
gapminder[gapminder$lifeExp > 80, ]
#Extract the first row, and the fourth and fifth columns (lifeExp and gdpPercap).
gapminder[1, c(4, 5)]
#Advanced: extract rows that contain information for the years 2002 and 2007
gapminder[gapminder$year %in% c(2002, 2007), ]

# Challenge 8
#1. gapminder[1:20] returns an error because gapminder is a 2D object with rows and columns.
#2.
gapminder_small <- rbind(gapminder[1:9, ], gapminder[19:23, ])
```

## Creating Publication-Quality Graphics

```{r}
library("ggplot2")

# Challenge 1
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point()

# Challenge 2
ggplot(data = gapminder, aes(x = year, y = lifeExp, color = continent)) + geom_point()

# Challenge 3
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) +
  geom_point() +
  geom_line(aes(color=continent)) # The line layer is now on top.

# Challenge 4a
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point(color = "red", size = 2) + scale_x_log10() + geom_smooth(method="lm", size=1.5)

# Challenge 4b
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point(color = "red", size = 2, aes(pch = continent)) + scale_x_log10() + geom_smooth(method="lm", size=1.5)

# Challenge 5
ggplot(data = gapminder, aes(x = gdpPercap, fill = continent)) +
  geom_density(alpha = 0.5) + facet_wrap(~year) + scale_x_log10()

```

